Supplementary material for ``An Efficient Forecasting Approach to Reduce Boundary Effects in Real-Time Time-Frequency Analysis''

### 2020-09-09 Notes/Supplementary material/SupMat.tex
proof

\begin{proof}
It follows from the signal model~\eqref{I-eq:sum.sine} that the matrices $\bS^{(0)}$ and $\bS^{(1)}$ take the following form:
\begin{align}
\nonumber
\bS^{(a)}[m,m'] &= \sigma^2\delta_{(m+a)\,,m'}+\sum_{j,j'=1}^J\dfrac{\Omega_j\Omega_{j'}}{K}\sum_{k=0}^{K-1} \cos\left(2\pi \frac{f_j}{\fs}(N_0+m+a+k)+\varphi_j\right)\cos\left(2\pi \frac{f_{j'}}{\fs}(N_0+m'+k)+\varphi_{j'}\right) \\
\nonumber
&= \sigma^2\delta_{(m+a)\,,m'}+\sum_{j=1}^J\dfrac{\Omega_j^2}{2K}\sum_{k=0}^{K-1} \cos\left(2\pi \frac{f_j}{\fs}(m+a-m')\right) + \cos\left(2\pi \frac{f_j}{\fs}(2k+m+a+m'+2N_0)\right)\\
\nonumber
&= \sigma^2\delta_{(m+a)\,,m'}+\sum_{j=1}^J\Bigg( \dfrac{\Omega_j^2}2\cos\left(2\pi \frac{f_j}{\fs}(m+a-m')\right) + \dfrac{\Omega_j^2}{2K}\underbrace{\sum_{k=0}^{K-1}\cos\left(2\pi \frac{f_j}{\fs}(2k+m+a+m'+2N_0)\right)}_{=0\ \mathrm{because}\ \frac{f_j}\fs=\frac{p'_j}K } \Bigg)\\
&= \sigma^2\delta_{(m+a)\,,m'}+\sum_{j=1}^J\dfrac{\Omega_j^2}2\cos\left(2\pi \frac{f_j}{\fs}(m+a-m')\right)\ .
\label{eq:S1}
\end{align}
Thus, $\bS^{(0)}$ is a circulant matrix, and is therefore diagonalizable in the Fourier basis:
\[
\bS^{(0)} = \bU\bLambda^{(0)}\bU^*\ ,
\]
where $\bU[m,m']=\frac1{\sqrt{M}}e^{-2\ii\pi mm'/M}$ and $\bLambda^{(0)} = \mathrm{diag}(\lambda_0^{(0)},\dots,\lambda_{M-1}^{(0)})$ with
\begin{align*}
\lambda_m^{(0)} &= \sigma^2 + \sum_{j=1}^J\dfrac{\Omega_j^2}2\sum_{q=0}^{M-1} \cos\left(2\pi\frac{f_j}{\fs} q\right) e^{-2\ii\pi qm/M} \\
&= \sigma^2 + \dfrac{M}{4}\sum_{j=1}^J\Omega_j^2(\delta_{m,p_j} + \delta_{m,M-p_j})\ .
\end{align*}
Therefore,
\begin{align*}
{\bS^{(0)}}^\inv  &= \bU{\bLambda^{(0)}}^\inv\bU^*\,,
\end{align*}
which leads to
\begin{equation}
{\bS^{(0)}}^\inv[m,m']  = \dfrac1{\sigma^2}\delta_{m,m'}-\dfrac{2}{M\sigma^2}\sum_{j=1}^J\dfrac{1}{1+\frac{4\sigma^2}{M\Omega_j^2}}\cos\left(2\pi p_j \dfrac{m-m'}{M}\right)\,.
\label{eq:S0.inv.2}
\end{equation}
Directly, we have:
\begin{align*}
\left|{\bS^{(0)}}^\inv[m,m']\right|  &= \dfrac1{\sigma^2}\left|\delta_{m,m'}-\dfrac{2}{M}\sum_{j=1}^J\dfrac{1}{1+\frac{4\sigma^2}{M\Omega_j^2}}\cos\left(2\pi p_j \dfrac{m-m'}{M}\right)\right| \\
&\leq \dfrac1{\sigma^2}\left( 1+\dfrac{2}{M}\sum_{j=1}^J\dfrac{1}{1+\frac{4\sigma^2}{M\Omega_j^2}} \right) \leq \dfrac1{\sigma^2}\left( 1+\dfrac{2J}{M} \right)\ .
\end{align*}
Thus, 
\[
\left\|{\bS^{(0)}}^\inv\right\|_{\max}\leq\dfrac1{\sigma^2}\left( 1+\dfrac{2J}{M} \right)\ .
\]
Furthermore, combining equations~\eqref{eq:S1} and~\eqref{eq:S0.inv.2}, we have
\begin{align*}
\bA_0[m,m']  &= \sum_{q=0}^{M-1} {\bS^{(1)}}[m,q]{\bS^{(0)}}^\inv[q,m'] \\
&= \delta_{m+1,m'} + \dfrac2M\sum_{j=1}^J\dfrac{1}{1+\frac{4\sigma^2}{M\Omega_j^2}}\cos\left(2\pi p_j\frac{m'}{M}\right)\delta_{m+1,M}\,.
\end{align*}
Directly, we have:
\begin{align*}
\left|\bA_0[m,m']\right| &\leq \left\{
\begin{array}{ll}
1 & \mbox{if}\quad m<M-1\,, \\
\dfrac{2}{M}\sum_{j=1}^J\dfrac{1}{1+\frac{4\sigma^2}{M\Omega_j^2}} & \mbox{if}\quad m=M-1\ . \\
\end{array}
\right.
\end{align*}
Thus, 
\[
\left\|\bA_0\right\|_{\max}\leq \max\left(1,\frac{2J}{M}\right)\ .
\]
\end{proof}

@@

Concerning higher-order moments, let $T\geq 3$ denote the order of the moment defined by $\EE\{\prod_{\theta=1}^T\bg[r_\theta]\}$. Thus,
\begin{align}
\nonumber
\EE\left\{\prod_{\theta=1}^T\bg[r_\theta]\right\}& = \EE\left\{\prod_{\theta=1}^T \left(\sigma\bg_1[r_\theta] + \sigma^2\bg_2[r_\theta] \right)\right\}\\
\nonumber
&=\dfrac1{K^T}\EE\left\{\prod_{\theta=1}^T \left(\sum_{k=0}^{K-1}\rho_{\theta,k}\left(\bw[k+m'_{r_\theta}],\bw[k+m_{r_\theta}+a_{r_\theta}]\right) \right)\right\}\\ 
&=\dfrac1{K^T}\sum_{k_1=0}^{K-1}\cdots\sum_{k_T=0}^{K-1}\EE\left\{\prod_{\theta=1}^T \rho_{\theta,k_\theta}\left(\bw[k_\theta+m'_{r_\theta}],\bw[k_\theta+m_{r_\theta}+a_{r_\theta}]\right) \right\} \,,
\label{eq:multisum}
\end{align}
where
\[
\rho_{\theta,k}(u,v)=\sigma \bz_k[m_{r_\theta}+a_{r_\theta}]u + \sigma \bz_k[m'_{r_\theta}]v + \sigma^2uv - \delta_{m_{r_\theta}+a_{r_\theta},\,m'_{r_\theta}}\ .
\]
Thus,
\begin{equation*}
\left|\EE\left\{\prod_{\theta=1}^T\bg[r_\theta]\right\}\right| \leq \dfrac{\nu_K}{K^T} C_T\ ,
\end{equation*}
where
\[
C_T = \max_{(k_1,\ldots,k_T)} \left|\EE\left\{\prod_{\theta=1}^T \rho_{\theta,k_\theta}(\bw[k_\theta+m'_{r_\theta}],\bw[k_\theta+m_{r_\theta}+a_{r_\theta}]) \right\} \right| \,,
\]
and $\nu_K$ is the number of non-zero terms in the sum~\eqref{eq:multisum}. Note that $C_T$ is independent of $K$ (but depends on $\bz$ and $\sigma$). The behavior of the order $T$ moment in function of $K$ is therefore only determined by the ratio $\dfrac{\nu_K}{K^T}$.

Let us bound $\nu_K$. Fix $k_1$. For each of the other indexes of summation $k_\theta$ ($\theta\in\{2,\ldots,T\}$), there are four values that make $\rho_{\theta,k_\theta}(\bw[k_\theta+m'_{r_\theta}],\bw[k_\theta+m_{r_\theta}+a_{r_\theta}])$ not independent of $\rho_{1,k_1}(\bw[k_1+m'_{r_1}],\bw[k_1+m_{r_1}+a_{r_1}])$. Indeed, since $\bw$ is a white noise these quantities are independent except when $k_\theta$ is such that:
\begin{align*}
k_\theta &= k_1+m'_{r_1}-m'_{r_\theta} \\
k_\theta &= k_1+m_{r_1}+a_{r_1} - m'_{r_\theta}\\
k_\theta &= k_1+m'_{r_1} - m_{r_\theta}- a_{r_\theta} \\
k_\theta &= k_1+m_{r_1}+a_{r_1} - m_{r_\theta}- a_{r_\theta}\ .
\end{align*}
Consequently, for each value of $k_1$ there exist at least $(K-4)^{T-1}$ combinations of $(k_2,\ldots,k_T)$ such that the correlation term given by $\EE\left\{\prod_{\theta=1}^T \rho_{\theta,k_\theta}(\bw[k_\theta+m'_{r_\theta}],\bw[k_\theta+m_{r_\theta}+a_{r_\theta}]) \right\}$ is zero. Because $T\geq 3$, we develop similar arguments on $k_2$ and $k_3$ to determine other cases where this correlation term vanishes. We subtract these cases to $K^T$, the total number of combinations of $(k_1,\ldots,k_T)$ to obtain the following maximum bound on the number of non-zero terms in the sum~\eqref{eq:multisum}:
\[
\nu_K\leq K^T - 3K(K-4)^{T-1} + 3K(K-4)(K-8)^{T-2}- K(K-4)(K-8)(K-12)^{T-3}\ .
\]
Thus,
\begin{align*}
\left|\EE\left\{\prod_{\theta=1}^T\bg[r_\theta]\right\}\right| &\leq C_T \dfrac{K^T - 3K(K-4)^{T-1} + 3K(K-4)(K-8)^{T-2}- K(K-4)(K-8)(K-12)^{T-3}}{K^T} \\
&\leq C_T \left( 1 - 3\left(1-\frac4K\right)^{T-1} + 3\left(1-\frac4K\right)\left(1-\frac8K\right)^{T-2}- \left(1-\frac4K\right)\left(1-\frac8K\right)\left(1-\frac{12}{K}\right)^{T-3} \right) \\
&\leq C_T \left( 1 - 3 +\frac{12(T-1)}{K} + 3-\frac{12}K-\frac{24(T-2)}{K} -1 + \frac{4}K + \frac{8}K +\frac{12(T-3)}{K} + o\left(\dfrac1K\right)\right) \\
&\leq C_T\ o\left(\dfrac1K\right)\ . \\
\end{align*}
Therefore,
\begin{equation*}
\left|\EE\left\{\prod_{\theta=1}^T\bg[r_\theta]\right\}\right| = o\left(\dfrac1K\right)\,,\quad\forall T\geq 3\ .
\end{equation*}
\end{proof}

@@

\left|\left.\dfrac{\partial^2 f_m^{(\ell)}}{\partial \bg[r]\partial \bg[r']}\right\vert_{\bg=\bzero}\right| \leq \dfrac{d_{2,\bz,M,\ell}}{\sigma^4}\ .

>>>>

\begin{proof}
It follows from the signal model~\eqref{I-eq:sum.sine} that the matrices $\bS^{(0)}$ and $\bS^{(1)}$ take the following form:
\begin{align}
\nonumber
\bS^{(a)}[m,m'] &= \sigma^2\delta_{(m+a)\,,m'}+\sum_{j,j'=1}^J\dfrac{\Omega_j\Omega_{j'}}{K}\sum_{k=0}^{K-1} \cos\left(2\pi \frac{f_j}{\fs}(N_0+m+a+k)+\varphi_j\right)\cos\left(2\pi \frac{f_{j'}}{\fs}(N_0+m'+k)+\varphi_{j'}\right) \\
\nonumber
&= \sigma^2\delta_{(m+a)\,,m'}+\sum_{j=1}^J\dfrac{\Omega_j^2}{2K}\sum_{k=0}^{K-1} \cos\left(2\pi \frac{f_j}{\fs}(m+a-m')\right) + \cos\left(2\pi \frac{f_j}{\fs}(2k+m+a+m'+2N_0)\right)\\
\nonumber
&= \sigma^2\delta_{(m+a)\,,m'}+\sum_{j=1}^J\Bigg( \dfrac{\Omega_j^2}2\cos\left(2\pi \frac{f_j}{\fs}(m+a-m')\right) + \dfrac{\Omega_j^2}{2K}\underbrace{\sum_{k=0}^{K-1}\cos\left(2\pi \frac{f_j}{\fs}(2k+m+a+m'+2N_0)\right)}_{=0\ \mathrm{because}\ \frac{f_j}\fs=\frac{p'_j}K } \Bigg)\\
&= \sigma^2\delta_{(m+a)\,,m'}+\sum_{j=1}^J\dfrac{\Omega_j^2}2\cos\left(2\pi \frac{f_j}{\fs}(m+a-m')\right)\ .
\label{eq:S1}
\end{align}
Thus, $\bS^{(0)}$ is a circulant matrix, and is therefore diagonalizable in the Fourier basis:
\[
\bS^{(0)} = \bU\bLambda^{(0)}\bU^*\ ,
\]
where $\bU[m,m']=\frac1{\sqrt{M}}e^{-2\ii\pi mm'/M}$ and $\bLambda^{(0)} = \mathrm{diag}(\lambda_0^{(0)},\dots,\lambda_{M-1}^{(0)})$ with
\begin{align*}
\lambda_m^{(0)} &= \sigma^2 + \sum_{j=1}^J\dfrac{\Omega_j^2}2\sum_{q=0}^{M-1} \cos\left(2\pi\frac{f_j}{\fs} q\right) e^{-2\ii\pi qm/M} \\
&= \sigma^2 + \dfrac{M}{4}\sum_{j=1}^J\Omega_j^2(\delta_{m,p_j} + \delta_{m,M-p_j})\ .
\end{align*}
Therefore,
\begin{align*}
{\bS^{(0)}}^\inv  &= \bU{\bLambda^{(0)}}^\inv\bU^*\,,
\end{align*}
which leads to
\begin{equation}
{\bS^{(0)}}^\inv[m,m']  = \dfrac1{\sigma^2}\delta_{m,m'}-\dfrac{2}{M\sigma^2}\sum_{j=1}^J\dfrac{1}{1+\frac{4\sigma^2}{M\Omega_j^2}}\cos\left(2\pi p_j \dfrac{m-m'}{M}\right)\,.
\label{eq:S0.inv.2}
\end{equation}
Directly, we have:
\begin{align*}
\left|{\bS^{(0)}}^\inv[m,m']\right|  &= \dfrac1{\sigma^2}\left|\delta_{m,m'}-\dfrac{2}{M}\sum_{j=1}^J\dfrac{1}{1+\frac{4\sigma^2}{M\Omega_j^2}}\cos\left(2\pi p_j \dfrac{m-m'}{M}\right)\right| \\
&\leq \dfrac1{\sigma^2}\left( 1+\dfrac{2}{M}\sum_{j=1}^J\dfrac{1}{1+\frac{4\sigma^2}{M\Omega_j^2}} \right) \leq \dfrac1{\sigma^2}\left( 1+\dfrac{2J}{M} \right)\ .
\end{align*}
Thus, 
\[
\left\|{\bS^{(0)}}^\inv\right\|_{\max}\leq\dfrac1{\sigma^2}\left( 1+\dfrac{2J}{M} \right)\ .
\]
Furthermore, combining equations~\eqref{eq:S1} and~\eqref{eq:S0.inv.2}, we have
\begin{align*}
\bA_0[m,m']  &= \sum_{q=0}^{M-1} {\bS^{(1)}}[m,q]{\bS^{(0)}}^\inv[q,m'] \\
&= \delta_{m+1,m'} + \dfrac2M\sum_{j=1}^J\dfrac{1}{1+\frac{4\sigma^2}{M\Omega_j^2}}\cos\left(2\pi p_j\frac{m'}{M}\right)\delta_{m+1,M}\,.
\end{align*}
Directly, we have:
\begin{align*}
\left|\bA_0[m,m']\right| &\leq \left\{
\begin{array}{ll}
1 & \mbox{if}\quad m<M-1\,, \\
\dfrac{2}{M}\sum_{j=1}^J\dfrac{1}{1+\frac{4\sigma^2}{M\Omega_j^2}} & \mbox{if}\quad m=M-1\ . \\
\end{array}
\right.
\end{align*}
Thus, 
\[
\left\|\bA_0\right\|_{\max}\leq \max\left(1,\frac{2J}{M}\right)\ .\qedhere
\]
\end{proof}

@@

Concerning higher-order moments, let $T\geq 3$ denote the order of the moment defined by $\EE\{\prod_{\theta=1}^T\bg[r_\theta]\}$. Thus,
\begin{align}
\nonumber
\EE\left\{\prod_{\theta=1}^T\bg[r_\theta]\right\}& = \EE\left\{\prod_{\theta=1}^T \left(\sigma\bg_1[r_\theta] + \sigma^2\bg_2[r_\theta] \right)\right\}\\
\nonumber
&=\dfrac1{K^T}\EE\left\{\prod_{\theta=1}^T \left(\sum_{k=0}^{K-1}\rho_{\theta,k}\left(\bw[k+m'_{r_\theta}],\bw[k+m_{r_\theta}+a_{r_\theta}]\right) \right)\right\}\\ 
&=\dfrac1{K^T}\sum_{k_1=0}^{K-1}\cdots\sum_{k_T=0}^{K-1}\EE\left\{\prod_{\theta=1}^T \rho_{\theta,k_\theta}\left(\bw[k_\theta+m'_{r_\theta}],\bw[k_\theta+m_{r_\theta}+a_{r_\theta}]\right) \right\} \,,
\label{eq:multisum}
\end{align}
where
\[
\rho_{\theta,k}(u,v)=\sigma \bz_k[m_{r_\theta}+a_{r_\theta}]u + \sigma \bz_k[m'_{r_\theta}]v + \sigma^2uv - \delta_{m_{r_\theta}+a_{r_\theta},\,m'_{r_\theta}}\ .
\]
Thus,
\begin{equation*}
\left|\EE\left\{\prod_{\theta=1}^T\bg[r_\theta]\right\}\right| \leq \dfrac{\nu_K}{K^T} C_T\ ,
\end{equation*}
where
\[
C_T = \max_{(k_1,\ldots,k_T)} \left|\EE\left\{\prod_{\theta=1}^T \rho_{\theta,k_\theta}(\bw[k_\theta+m'_{r_\theta}],\bw[k_\theta+m_{r_\theta}+a_{r_\theta}]) \right\} \right| \,,
\]
and $\nu_K$ is the number of nonzero terms in the sum~\eqref{eq:multisum}. Note that $C_T$ is independent of $K$ (but depends on $\bz$ and $\sigma$). The behavior of the order $T$ moment in function of $K$ is therefore only determined by the ratio $\dfrac{\nu_K}{K^T}$.

Let us bound $\nu_K$. Fix $k_1$. For each of the other indexes of summation $k_\theta$ ($\theta\in\{2,\ldots,T\}$), there are four values that make $\rho_{\theta,k_\theta}(\bw[k_\theta+m'_{r_\theta}],\bw[k_\theta+m_{r_\theta}+a_{r_\theta}])$ not independent of $\rho_{1,k_1}(\bw[k_1+m'_{r_1}],\bw[k_1+m_{r_1}+a_{r_1}])$. Indeed, since $\bw$ is a white noise these quantities are independent except when $k_\theta$ is such that:
\begin{align*}
k_\theta &= k_1+m'_{r_1}-m'_{r_\theta} \\
k_\theta &= k_1+m_{r_1}+a_{r_1} - m'_{r_\theta}\\
k_\theta &= k_1+m'_{r_1} - m_{r_\theta}- a_{r_\theta} \\
k_\theta &= k_1+m_{r_1}+a_{r_1} - m_{r_\theta}- a_{r_\theta}\ .
\end{align*}
Consequently, for each value of $k_1$ there exist at least $(K-4)^{T-1}$ combinations of $(k_2,\ldots,k_T)$ where we have
\begin{align*}
\EE\left\{\prod_{\theta=1}^T \rho_{\theta,k_\theta}(\bw[k_\theta+m'_{r_\theta}],\bw[k_\theta+m_{r_\theta}+a_{r_\theta}]) \right\} &= \EE\left\{ \rho_{1,k_1}(\bw[k_1+m'_{r_1}],\bw[k_1+m_{r_1}+a_{r_1}]) \right\} \\
&\hspace{20pt}\times\EE\left\{\prod_{\theta=2}^T \rho_{\theta,k_\theta}(\bw[k_\theta+m'_{r_\theta}],\bw[k_\theta+m_{r_\theta}+a_{r_\theta}]) \right\}
& = 0\,,
\end{align*}
because $\EE\left\{ \rho_{1,k_1}(\bw[k_1+m'_{r_1}],\bw[k_1+m_{r_1}+a_{r_1}]) \right\}=0$. Therefore, at least $K(K-4)^{T-1}$ of the sum~\eqref{eq:multisum} are zero. Because $T\geq 3$, we develop similar arguments on $k_2$ and $k_3$ to determine other cases where this correlation term vanishes. We subtract these cases to $K^T$, the total number of combinations of $(k_1,\ldots,k_T)$ to obtain the following maximum bound on the number of nonzero terms in the sum~\eqref{eq:multisum}:
\[
\nu_K\leq K^T - 3K(K-4)^{T-1} + 3K(K-4)(K-8)^{T-2}- K(K-4)(K-8)(K-12)^{T-3}\ .
\]
Thus,
\begin{align*}
\left|\EE\left\{\prod_{\theta=1}^T\bg[r_\theta]\right\}\right| &\leq C_T \dfrac{K^T - 3K(K-4)^{T-1} + 3K(K-4)(K-8)^{T-2}- K(K-4)(K-8)(K-12)^{T-3}}{K^T} \\
&\leq C_T \left( 1 - 3\left(1-\frac4K\right)^{T-1} + 3\left(1-\frac4K\right)\left(1-\frac8K\right)^{T-2}- \left(1-\frac4K\right)\left(1-\frac8K\right)\left(1-\frac{12}{K}\right)^{T-3} \right) \\
&\leq C_T \left( 1 - 3 +\frac{12(T-1)}{K} + 3-\frac{12}K-\frac{24(T-2)}{K} -1 + \frac{4}K + \frac{8}K +\frac{12(T-3)}{K} + o\left(\dfrac1K\right)\right) \\
&\leq C_T\ o\left(\dfrac1K\right)\ . \\
\end{align*}
Therefore,
\begin{equation*}
\left|\EE\left\{\prod_{\theta=1}^T\bg[r_\theta]\right\}\right| = o\left(\dfrac1K\right)\,,\quad\forall T\geq 3\ .\qedhere
\end{equation*}
\end{proof}

@@

\left|\left.\dfrac{\partial^2 f_m^{(\ell)}}{\partial \bg[r]\partial \bg[r']}\right\vert_{\bg=\bzero}\right| \leq \dfrac{d_{2,\bz,M,\ell}}{\sigma^4}\ .\qedhere
